---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(data.table)
library(readtext)
library(lubridate)
library(ggthemes)
```

#Getting data

Data from:
http://www.planet4589.org/space/lvdb/list2.html

Name of the fields: http://www.planet4589.org/space/lvdb/launchcols.html


```{r}

dataG <- read.fwf("http://www.planet4589.org/space/lvdb/lis/G.lis", widths = c(15,11,21,25,7,21,25,25,25,10,9,17,7,2,5,2,13,13,5,25,25,21,21),strip.white=TRUE,col.names=c("ref","id","Launchdate","LV_Type","LV_Variant","Flight_ID","Flight","Mission","flightcode", "Platform", "Launch_Site","Launch_Pad","Apogee","apoflag","Range","rangeflag","Destination","Launch_Agency","Launch_Code","Group","Category","Citation1","Citation2"),na.strings="-")
dataG$success <- 
dataO <- read.fwf("http://www.planet4589.org/space/lvdb/lis/O.lis", widths = c(15,11,21,25,7,21,25,25,25,10,9,17,7,2,5,2,13,13,5,25,25,21,21),strip.white=TRUE,col.names=c("ref","id","Launchdate","LV_Type","LV_Variant","Flight_ID","Flight","Mission","flightcode", "Platform", "Launch_Site","Launch_Pad","Apogee","apoflag","Range","rangeflag","Destination","Launch_Agency","Launch_Code","Group","Category","Citation1","Citation2"),na.strings="-")
dataF <- read.fwf("http://www.planet4589.org/space/lvdb/lis/F.lis", widths = c(15,11,21,25,7,21,25,25,25,10,9,17,7,2,5,2,13,13,5,25,25,21,21),strip.white=TRUE,col.names=c("ref","id","Launchdate","LV_Type","LV_Variant","Flight_ID","Flight","Mission","flightcode", "Platform", "Launch_Site","Launch_Pad","Apogee","apoflag","Range","rangeflag","Destination","Launch_Agency","Launch_Code","Group","Category","Citation1","Citation2"),na.strings="-")
```



```{r}
data <- rbind(dataG, dataF,dataO)
```

```{r}
150+344+5267
```

```{r}
 data$Launch_Agency <- gsub("KTRAS/RVSNR", "KTRAS", data$Launch_Agency)
```



Organisation names
http://www.planet4589.org/space/lvdb/sdb/Orgs 

Documentation:http://www.planet4589.org/space/lvdb/orgs.html

```{r}
org <- read.fwf('http://www.planet4589.org/space/lvdb/sdb/Orgs',
               widths=c(9,9,7,17,2,13,13,18,81,53,13,11,8,13,17,61,100),skip = 59,col.names=c('code','ucode','parent_code','phase_type','class','start_dates','stop_dates','sname','fullname','location','long','lat','error','parent_org_phase','sname_eng','fullname_eng','region'),strip.white=TRUE,na.strings='-')
```

```{r}
org
```


```{r}
total = merge(data, org, by.x = 'Launch_Agency', by.y = 'code')
```

```{r}
anti_join(data,total)
#not included in total dataset
```

```{r}
dataO %>% filter(Launch_Agency == "SIS/ELUS")
```

#Cleaning data

```{r}
unique(colnames(total))
```

```{r}
total_small <- total %>% select(Launch_Agency, ref, Launchdate, LV_Type, Flight_ID, Flight, Mission, Launch_Site, Launch_Pad, Launch_Code, Category, ucode,parent_code, phase_type, class, sname, fullname,location, sname_eng, fullname_eng)
```


`
```{r}
#cleaning unknown agencies
total_small <-  total_small %>% filter(Launch_Agency !="UNK")
total_small <- total_small %>% mutate(Year = str_extract(ref, "[0-9][0-9][0-9][0-9]"))
```


```{r}

ggplot(total_small, aes(x=Year,fill=parent_code)) +geom_bar()
```

# Private or statal companies?

```{r}
state <- c("USAF", "U.S. Army Ballistic Missile Agency", "US Air Force Materiel Command","US Air Force Systems Command","United States Army","United States Navy","Agencia Espacial Brasileira","Zhongguo yunzaihuojian jishu yanjiu yuan","Centre Nationale d'Etudes Spatiales","Centro Richerchi Aerospaziale","Centre Europeene de Construct. de Lanceurs et d'Engins Spatiaux","European Space Agency","Hangtian","Federal'noe kosmicheskoe agentstvo Rossii","Sazman Fizaii Iran","Israeli Space Agency","Ucyu Koku Kenkyujyo","Indian Space Research","Ucyu Koku Kenkyu Kaihatsu Kikou","Han-guk Hanggong Uju","Joseon ujugong gangisul","GKNPTs im. M.V. Khrunichev","Kosmichesikiye voyska Rossii","Ministerstvo Voenniy","Gugga Uju Gaebalgug","National Aeronautics and Space Administration","Ucyu Kaihatsu Jigyoudan","Royal Aircraft Establishment","Rossiskoe aviatsionno","Rakentiye Voiska Strategicheskogo Naznacheniye","Shanghai hangtian jishu yanjiuyuan","Space International Services","Upravleniye Nachalnika Kosmicheskikh Sredstv","Voenno-Kosmicheskiye","Voenno-Morskoy ","Voisk Vosdushno-Kosmicheskoi Oborona
MKK Kosmotras)")

private <- c("Arianespace","Boeing Launch Services (?)","EER Systems Inc. Aerospace Div","Energia Logistics Ltd","Eurockot Launch Services GmbH","General Dynamics Commercial Launch","International Launch Services","Beijing lanjian kongjian keji youxian gongsi","Lockheed Martin Astronautics","Lockheed Missiles and Space Co.","McDonnell Douglas Space Systems Corp.","Mitsubishi Jukogyo Kabushiki-kaisha","Martin Marietta","Orbital ATK","Orbital Sciences Corp.","ZAO Puskovie Uslugi","Rocket Lab USA","Rocket Systems Corporation","Sea Launch Limited Partnership","SpaceX","Starsem","United Launch Alliance/Boeing")

```

```{r}
write.csv(total_small %>% group_by(fullname) %>% count(),"agencies.csv")
```

```{r}
agencies <- read.csv("agencies.csv", sep = ";")
space <- merge(total_small, agencies, by = "fullname")
```

```{r}
space <- space %>% mutate(Success = ifelse(Launch_Code == "OF","F","O")) 
space <- space %>% select(fullname, Launchdate, Launch_Agency, Mission, parent_code, location, fullname_eng, Year, state, Success)
```

```{r}
space$parent_code <- gsub(c("SU|RU"),"SU/RU",space$parent_code)
space$parent_code <- gsub(c("CYM|UK"),"CYM/UK",space$parent_code)
```


# Visualisation

```{r}
space %>% group_by(state) %>% count(state) %>% mutate(n/(953+4801))
```

```{r}

ggplot(space, aes(x=Year, fill=state))+geom_bar(aes(alpha=Success))+ 
  scale_alpha_discrete(range=c(0.5, 1))+ theme(axis.text.x = element_text(angle = 90))
```

83% of the space battle has been state financially support, mainly by US and USSR.

Mid 80s marks the starting point for private companies. Arianespace came into scene and the presence of private space explorations projects have been increasing since then.

```{r}
ggplot(space %>% filter(Year<=1984) %>% arrange(desc(Year)), aes(x=Year, fill=parent_code))+geom_bar()+ theme(axis.text.x = element_text(angle = 90))

```

First three decades, the US first and the Soviet Union since mid-65 dominates the battle, with some but little European and Asian influence. 

```{r}
ggplot(space %>% arrange(desc(Year)), aes(x=Year, fill=parent_code))+geom_bar(aes(alpha=state))+ theme(axis.text.x = element_text(angle = 90))+ 
  scale_alpha_discrete(range=c(0.4, 1))
```

The end of the decade of the 80s the space projects public funded dropped, and it only recovered in the early 2000 when China increases its investement. In the last 10 years, the Chinese agency has been the only one with a significant increase in the rockets. 

```{r}
ggplot(space %>% filter(Year > 2003 & state == "state" & parent_code %in% c("CN", "SU/RU")) %>% group_by(parent_code, Year) %>% count(), aes(x=Year, y=n, group=parent_code))+geom_line(aes(color=parent_code))
```




```{r}
ggplot(space %>% filter(Year >=1984) %>%arrange(desc(Year)), aes(x=Year, fill=parent_code))+geom_bar(aes(alpha=state))+ theme(axis.text.x = element_text(angle = 90))+ 
  scale_alpha_discrete(range=c(1, .3))
```

As for the number of project privately supported, these have been increasing since the late 80s, with a reducton 2001-2005 and 2010. The US and France (Europe?) dominates these battle, although the Russian adn Japanese presence is remarkable. 

```{r}
write.csv(space, "spacedata.csv")
```

